---
title: "GSA 2021 Poster Code"
author: "Peter Sun"
date: "October 22, 2021"
output:
  pdf_document:
    number_sections: true
header-includes:
- \usepackage{enumitem}
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
---

```{r setup, include=FALSE}
options(scipen=999)
knitr::opts_chunk$set(echo = TRUE)
options(knitr.kable.NA = "") # omit NAs in Kable
```

# Load Packages and Data

```{r message=F, error=F, warning=F}
library(tidyverse)
library(haven)
library(sjlabelled)
library(cobalt)
library(MatchThem)
library(mice)

# Avoid select clashes
select <- dplyr::select
recode <- dplyr::recode
summarize <- dplyr::summarize

# Import RAND 2010 HRS Fat File (hd10f5f)
rand.10 <- read_dta(Sys.getenv("HRS_2010_FAT"), col_select = c(

  # Identifiers
  hhid, pn,
  
  # Live alone = Number of people living in the household
  tot_hh_members = ma098,
  
  # Eyesight Rating (Section C, Wave 10 / Year 2010)
  # 1 = excellent, 2 = very good, 3 = good, 4 = fair, 5 = poor
  # 6 = leagally blind, # 8 = DK, # 9 = RF
  vision = mc095,
  
  # Religious Participation (Section B, Wave 10 / Year 2010)
  religious = mb082,
  
)) %>%
  rename(
  vision = mc095,
  religious = mb082,
  tot_hh_members = ma098)

# Import Region Early V8.2
r <- read_dta(Sys.getenv("HRS_REGION"), col_select = c(
  
  # Identifiers
  hhid, pn,
  
  # HRS URBAN-RURAL CODE 2010 (BEALE 2013)
  beale2013_10
  
)) %>%
  rename(region = beale2013_10)

# Import Tracker Early 2018 V 2.0
t <- read_dta(Sys.getenv("HRS_TRACKER"), col_select = c(
  
  # Identifiers
  hhid, pn,
  
  # RESPONDENT WEIGHT FOR THE 2012 LEAVE BEHIND QNAIRE
  # See https://hrsonline.isr.umich.edu/sitedocs/wghtdoc.pdf
  nlbwgtr
  
)) %>%
  rename(lb_weight = nlbwgtr)

# Longitudinal RAND 1992_2018v1
rand <- read_dta(Sys.getenv("HRS_LONG"), col_select = c(
  
  # Identifiers
  hhid, pn,
  
  # Wave Information
  inw10, inw11,
  
  # Age at Exit Interview (Wave 10 / Year 2010)
  r10agey_e,
  
  # Gender (1 = male, 2 = female)
  ragender,
  
  # Race (1 = White/Caucasian, 2 = Black/African American, 3 = Other)
  raracem,
  
  # Marital Status (Wave 10 / Year 2010)
  # 2 = Married, spouse absent, 3 = partnered, 4 = separated, 
  # 5 = divorced, 6 = separated/divorced, 7 = widowed,
  # 8 = never married
  r10mstat,

  # Educational Degree 1 = < HS, 2 = GED, 3 = HS grad, 
  # 4 = some college, 5 = college and above
  raeduc,
  
  # Total Household Income (Wave 10 / Year 2010)
  h10iftot,
  
  # Self Report of Health (Wave 10 / Year 2010)
  r10shlt,
  
  # ADL (Wave 10 / Year 2010)
  r10adla,
  
  # IADL (Wave 10 / Year 2010)
  r10iadlza,
  
  # Poverty (Wave 10 / Year 2010)
  h10inpov,
  
  # Chronic health conditions (Wave 10 / Year 2010)
  # (high bp, diabetes, cancer, 
  # lung, heart, stroke, psych, arthritis)
  h1 = r10hibp, h2 = r10diab, h3 = r10cancr, h4 = r10lung, h5 = r10heart, 
  h6 = r10strok, h7 = r10psych, h8 = r10arthr,

  # Depression (Wave 10 / Year 2010)
  r10cesd,
  
  # Nursing Home (Wave 10 / Year 2010)
  r10nhmliv,
  
  # Medicaid (Wave 10 / Year 2010)
  r10govmd,
  
  # Retirement
  r10sayret
  
  )) %>%
  rename(age = r10agey_e,
    gender = ragender,
    race = raracem,
    married = r10mstat,
    degree = raeduc,
    income = h10iftot,
    self_health = r10shlt,
    adl = r10adla,
    iadl = r10iadlza,
    poverty = h10inpov,
    h1 = r10hibp, h2 = r10diab, h3 = r10cancr, 
    h4 = r10lung, h5 = r10heart, h6 = r10strok,
    h7 = r10psych, h8 = r10arthr,
    depression = r10cesd,
    nursing = r10nhmliv,
    medicaid = r10govmd,
    retired = r10sayret)

# Import Treatment, Outcome, and LBQ Survey Weights from 
# the RAND 2012 HRS Core Fat File (h12f3a)
rand.12 <- read_dta(Sys.getenv("HRS_2012_FAT"), col_select =c(
  
  # Identifiers
  hhid, pn,
  
  # HCBS Variables (NV051: attend congregate meal;
  # NV054: home delivered meals; NV057: transportation; NV060: case management
  # NV063: home services; NV068: caregiver services)
  # n6 = nv068, (including this is likely a mistake, see literature)
  n1 = nv051, n2 = nv054, n3 = nv057, n4 = nv060, n5 = nv063, 
  
  # Social engagement (volitional)
  d1 = nlb001c, d2 = nlb001d, d3 = nlb001e, d4 = nlb001f, d5 = nlb001g, d6 = nlb001h,
  d7 = nlb001i, d8 = nlb001k, d9 = nlb001l, d10 = nlb001m, d11 = nlb001n, d12 = nlb001o,
  d13 = nlb001p, d14 = nlb001q, d15 = nlb001r, d16 = nlb001s, d17 = nlb001t,
  
  # Completed leave-behind questionnaire
  lb_comp = nlbcomp
  
)) %>%
  rename(n1 = nv051, n2 = nv054, n3 = nv057, n4 = nv060, n5 = nv063,
        d1 = nlb001c, d2 = nlb001d, d3 = nlb001e, d4 = nlb001f, d5 = nlb001g, 
        d6 = nlb001h, d7 = nlb001i, d8 = nlb001k, d9 = nlb001l, 
        d10 = nlb001m, d11 = nlb001n, d12 = nlb001o, d13 = nlb001p, 
        d14 = nlb001q, d15 = nlb001r, d16 = nlb001s, d17 = nlb001t,
        lb_comp = nlbcomp)
```

\newpage
# Merge HRS Datasets and Clear Stata Formatting

```{r}
df <- rand.12 %>%
  left_join(t, by = c("hhid", "pn")) %>%
  left_join(r, by = c("hhid", "pn")) %>%
  left_join(rand, by = c("hhid", "pn")) %>%
  left_join(rand.10, by = c("hhid", "pn")) %>%
  haven::zap_formats() %>%
  sjlabelled::remove_all_labels() %>%
  as_tibble()
```

\newpage
# Recode Variables

```{r}
df.recoded <- df %>%
  
  # Recode marital status (0 = not married, 1 = married)
  # Marital status in wave 11 (2012), (1 = Married, 2 = Married, spouse absent, 
  # 3 = partnered, 4 = separated, 
  # 5 = divorced, 6 = separated/divorced, 
  # 7 = widowed, 8 = never married, else NAs)
  mutate(married = ifelse(married == 1, 1,
    ifelse(married %in% 2:8, 0, NA))) %>%
  
  # Live alone (1 = live alone, 0 = does not live alone or at least two people 
  # in the household, NA = NA)) (as a social isolation proxy)
  mutate(live_alone = ifelse(tot_hh_members == 0, 1, 
    ifelse(tot_hh_members >= 1, 0, NA))) %>%
  
  # Recode education (1 = < HS & GED, 2 = HS grad, 
  # 3 = some college & college and above)
  # Consider GED as less than HS based on 
  # https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4112190/
  mutate(degree = dplyr::recode(degree, 
    `1` = 1, `2` = 1, `3` = 2, `4` = 3, `5` = 3, 
    .default = NA_real_)) %>%
  
  # Recode HCBS Variables
  # HCBS = 1 if there are any 1s (services)
  # Remove sub_food, because it asks about "you or other family members"
  mutate_at(vars(n1:n5), 
    list(~dplyr::recode(., `1` = 1, `5` = 0, 
    .default = NA_real_))) %>% # 0 = "No", 1 = "Yes"
  mutate(hcbs_na = rowSums(is.na(
    dplyr::select_at(.,vars(n1:n5))))) %>% # Count HCBS NAs
  mutate(hcbs_sum = rowSums(dplyr::select(., n1, n2, n3, n4, n5), 
                            na.rm = TRUE)) %>% # Sum HCBSs
  mutate(hcbs = ifelse(
    hcbs_sum >= 1, 1, 0)) %>% # if at least one hcbs, then 1, otherwise 0
  mutate(hcbs = ifelse(
    hcbs_na == 5, NA, hcbs)) %>% # if 5 NAs, then NA, otherwise 0

  # Rural
  mutate(rural = recode(region, `1` = 0, `2` = 0, `3` = 1, 
    .default = NA_real_)) %>%
  
  # Chronic Health Conditions
  mutate(chronic = rowSums(
    ifelse(
      select(., h1, h2, h3, h4, h5, h6, h7, h8) == 1, 1, 0
      ), na.rm = TRUE)) %>% # count 1s from h1-h8
  
  # Vision (1 = excellent, 2 = very good, 3 = good, 4 = fair, 5 = poor, 
  # 6 = legally blind, 8 = don't know, 9 = refused)
  mutate(vision = recode(vision, `1` = 1, `2` = 1, `3` = 1, `4` = 0, 
    `5` = 0, `6` = 0, .default = NA_real_)) %>% 
  
  # Religious / How often attend religious service (1 = more than once a week, 
  # 2 = once a week, 3 = two or three times a week, 
  # 4 = one or more times a week, 5 = not at all, 8 = DK, 9 = RF)
  mutate(religious = recode(religious, `1` = 1, `2` = 1, `3` = 1, `4` = 1, 
    `5` = 0, .default = NA_real_)) %>%
  
  # Social Engagement
  mutate_at(vars(d1:d17), list(~dplyr::recode(., `1` = 1, `2` = 1, `3` = 1,
    `4` = 1, `5` = 0, `6` = 0, `7` = 0, 
    .default = NA_real_))) %>% # recode
  mutate(engagement_soc_sum = rowSums(
    dplyr::select_at(., vars(d1:d3, d5)), na.rm = TRUE)) %>%
  mutate(engagement_soc_na = rowSums(
    is.na(dplyr::select_at(.,vars(d1:d3, d5))))) %>% # count NAs
  mutate(engagement_soc_binary = ifelse(engagement_soc_sum >= 1, 1, 0)) %>%
  mutate(engagement_soc_sum = ifelse(
    engagement_soc_na > 2, NA, engagement_soc_sum)) %>% # NA if > 2 NAs
  mutate(engagement_soc_binary = 
           ifelse(engagement_soc_na > 2, NA, 
                  engagement_soc_binary)) %>% # set to NA if more than 2 NAs
  
  mutate(engagement_cog_sum = rowSums(dplyr::select_at(., 
    vars(d7:d11, d13:d15)), na.rm = TRUE)) %>%
  mutate(engagement_cog_na = rowSums(is.na(dplyr::select_at(.,
    vars(d7:d11, d13:d15))))) %>% # count NAs
  mutate(engagement_cog_binary = ifelse(
    engagement_cog_sum >= 1, 1, 0)) %>%
  mutate(engagement_cog_sum = ifelse(
    engagement_cog_na > 2, NA, 
    engagement_cog_sum)) %>% # set to NA if more than 2 NAs
  mutate(engagement_cog_binary = ifelse(
    engagement_cog_na > 2, NA, 
    engagement_cog_binary)) %>% # set to NA if more than 2 NAs
  
  mutate(engagement_phy_sum = rowSums(dplyr::select_at(., 
    vars(d12, d16, d17)), na.rm = TRUE)) %>%
  mutate(engagement_phy_na = rowSums(is.na(dplyr::select_at(.,
    vars(d12, d16, d17))))) %>% # count NAs
  mutate(engagement_phy_binary = ifelse(
    engagement_phy_sum >= 1, 1, 0)) %>%
  mutate(engagement_phy_sum = ifelse(engagement_phy_na > 2, NA, 
    engagement_phy_sum)) %>% # set to NA if more than 2 NAs
  mutate(engagement_phy_binary = ifelse(engagement_phy_na > 2, NA,
    engagement_phy_binary)) %>% # set to NA if more than 2 NAs

  # Collapse ADL and IADL
  # 0 = none, 1 = 1-2 ADLs, 2 = >=3 ADLs
  mutate(adl = recode(adl, `0` = 0, `1` = 1, `2` = 1, 
                      `3` = 2, `4` = 2, `5` = 2)) %>%
  mutate(iadl = recode(iadl, `0` = 0, `1` = 1, `2` = 1, 
                       `3` = 2, `4` = 2, `5`= 2)) %>%
  
  # Follow Muramatsu for income and wealth transformations
  mutate(income = ifelse(income == 0, income, log(income))) %>%
  
  # Retirement
  # Original: 0 = not retired, 1 = completely retired, 2 = partly retired, 
  # 3 = not working for pay or homemaker (change 3. question irrelevant to NA)
  mutate(retired = recode(retired, `0` = 0, `1` = 1, `2` = 2, 
                          .default = NA_real_)) %>%
  
  # Convert Categorical Variables to Factors
  mutate(race = factor(race),
         degree = factor(degree),
         self_health = factor(self_health),
         adl = factor(adl),
         iadl = factor(iadl),
         retired = factor(retired))
```

## Implement Sample Inclusion Criteria

```{r}
# Wave 11 (2012) Sample Size = 20554
nrow(df.recoded)
table(df.recoded$inw10)
table(df.recoded$inw11)

# 1. >= Age 51 = 18,554
df2 <- df.recoded %>%
  filter(age >= 51)
nrow(df2)

# 2. Not in nursing home = 18,293
df3 <- df2 %>%
  filter(nursing == 0)
nrow(df3)

# 3. Completed the 2012 LB (i.e., potentially have outcome variables) = 6,764
df4 <- df3 %>%
  filter(lb_comp %in% c(1, 2, 4))
nrow(df4)

# 4. Answered at least one HCBS question in Module 2 = 627
df5 <- df4 %>%
  filter(hcbs_na != 5)
nrow(df5)

# 5. No missing values in the dependent variables = 617
df6 <- df5 %>%
  filter(!is.na(engagement_soc_sum) & 
           !is.na(engagement_cog_sum) & 
           !is.na(engagement_phy_sum))
nrow(df6)
```

\newpage
# Model Specifications

```{r}
# Model Variables with Missing Data (n = 617)
df.final <- df6 %>%
  select(age, gender, race, married, degree, degree, income, self_health, adl, iadl, 
         depression, live_alone, hcbs, rural,
         chronic, poverty, religious, vision, 
         engagement_soc_sum, engagement_soc_binary, 
         engagement_phy_sum, engagement_phy_binary, engagement_cog_sum, 
         engagement_cog_binary, lb_weight, medicaid, retired)
nrow(df.final)

# Model Variables Without Missing Data (n = 562)
df.final.cca <- df.final %>%
  drop_na()
nrow(df.final.cca)

# % missing
1 - nrow(df.final.cca)/nrow(df.final)
```

\newpage
# Missing Data Analysis

```{r}
df.final %>%
  dplyr::summarize(across(everything(), ~sum(is.na(.)))) %>%
  rowid_to_column() %>%
  pivot_longer(-rowid) %>%
  mutate(pct = value/nrow(df.final)) %>%
  arrange(desc(pct)) %>%
  mutate(pct = scales::percent(pct, accuracy = 0.1))
```

\newpage
# Multiple Imputation

```{r}
# Multiple Imputation of Final Dataset
imp.m = 10
# df.final.imputes = mice(df.final, m = imp.m, maxit = 25, print = F)

# Load/Save Results
# save(list = c("df.final.imputes"), file = "data/df.final.imputes.Rdata")
load("data/df.final.imputes.Rdata")

# Inspect one of the complete data
df.final.imputed_data = complete(df.final.imputes, 5)
nrow(df.final.imputed_data)
```

\newpage
# Propensity Score Weighting

```{r}
# Define Propensity Score Model Formula
f.match <- as.formula(
  "hcbs ~ age + gender + race + married + degree + live_alone + 
  income + self_health + adl + iadl + rural + vision + poverty +
  chronic + religious + depression + lb_weight + medicaid + retired"
)

# Estimate Propensity Score Weights
# df.final.psw <- weightthem(
#   formula = f.match, 
#   datasets = df.final.imputes, 
#   approach = "within",
#   method = "gbm",
#   estimand = "ATT",
#   stop.method = "es.mean"
# )

# Load/Save Results
load("data/df.final.psw.Rdata")
# save(list = c("df.final.psw"), file = "data/df.final.psw.Rdata")

# Multiply the weights in each imputed dataset by the sampling weights
for(i in 1:imp.m) {
  df.final.psw$models[[i]]$weights <- 
    df.final.psw$models[[i]]$weights * df.final.psw$models[[i]]$covs$lb_weight
}

# Love Plot Labels
love.plot.names <- c(
  age = "Age", 
  gender_2 = "Female", 
  race_1 = "Race - White/Caucasian",
  race_2 = "Race - Black/African American",
  race_3 = "Race - Other", 
  married = "Married",
  degree_1 = "Education - Less than HS",
  degree_2 = "Education - HS Graduate",
  degree_3 = "Education - College and Above",
  live_alone = "Live Alone",
  income = "Income (Logged)",
  self_health_1 = "Self-Rated Health - Excellent",
  self_health_2 = "Self-Rated Health - Very Good",
  self_health_3 = "Self-Rated Health - Good",
  self_health_4 = "Self-Rated Health - Fair",
  self_health_5 = "Self-Rated Health - Poor",
  adl_0 = "ADLs - None",
  adl_1 = "ADLs - 1-2 ADLs",
  adl_2 = "ADLs - 3+ ADLs",
  iadl_0 = "IADLs - None",
  iadl_1 = "IADLs - 1-2 IADLs",
  iadl_2 = "IADLs - 3+ IADLs",
  rural = "Rural",
  vision = "Vision Score",
  poverty = "Poverty",
  chronic = "Chronic Conditions",
  religious = "Religious",
  depression = "Depression",
  lb_weight = "Survey Weight",
  medicaid = "Medicaid Coverage",
  retired_0 = "Retired - Not Retired",
  retired_1 = "Retired - Completely Retired",
  retired_2 = "Retired - Partly Retired"
  )

# Check Balance
lp <- love.plot(df.final.psw, 
  binary = "std", 
  stars = "std",
  drop.distance = TRUE,
  var.names = love.plot.names,
  shapes = c("circle filled", "circle"),
  threshold = c(m = .25),
  position = "top",
  title = NULL)
lp
```

\newpage
# Genetic Matching

```{r}
# # Genetic matching
# cl <- parallel::makeCluster(8)
# m1.gen <- matchthem(formula = f.match,
#           datasets = df.final.imputes,
#           approach = "within",
#           method = "genetic",
#           distance = "logit",
#           discard = "none",
#           estimand = "ATT",
#           pop.size = 1000,
#           cluster = cl)

# Load/Save Results
load("data/m1.gen.Rdata")
# save(list = c("m1.gen"), file = "data/m1.gen.Rdata")

# Check Balance
lp2 <- love.plot(m1.gen, 
  binary = "std", 
  stars = "std",
  var.names = love.plot.names,
  drop.distance = TRUE,
  shapes = c("circle filled", "circle"),
  threshold = c(m = .25),
  position = "top",
  title = "Covariate Balance After Genetic Matching")
lp2
ggsave("genetic.balance.png", dpi = 600, plot = lp2, 
       width = 10, height = 5, units = "in")
```

\newpage
# Outcome Analysis with Genetic Matching

```{r}
m1.gen.lm <- with(
  data = m1.gen, 
  exp = lm(engagement_soc_sum ~ hcbs + age + gender + 
           race + married + degree + live_alone + 
           income + self_health + adl + iadl + 
           rural + vision + poverty + chronic + 
           religious + depression + medicaid + retired)
)
m1.gen.lm.pooled <- pool(m1.gen.lm)
summary(m1.gen.lm.pooled)
```

\newpage
# Outcome Analysis with Propensity Score Weighting

```{r}
# Outcome Model
m1 <- with(
  data = df.final.psw, 
  exp = lm(engagement_soc_sum ~ hcbs + age + gender + 
           race + married + degree + live_alone + 
           income + self_health + adl + iadl + 
           rural + vision + poverty + chronic + 
           religious + depression + medicaid + retired)
)
m1.pooled <- pool(m1)
summary(m1.pooled)
```